\section{Skills}

Bogboa is a \emph{skills} based RPG engine.  You can think of a skill as both
a category for player abilities and rating of effectiveness when performing
those abilities in that category.  For example, the ability to mend the wounds
of others might fall within the skill \emph{healing}.  A character with a 
higher healing skill would naturally heal better than one with a lower rating
while a third character with no healing skill could not learn the ability in
the first place.

That last point is important.  In Bogboa, you don't wire abilities directly to
set classes.  You specifty a minimum skill level required to learn an ability
and any character who meets or exceeds that skill can obtain it.
\footnote{They might have to fetch a scroll from the bottom of a dungeon or 
pay a trainer hundreds of gold, however.}

This means that if you gave a melee class a tiny conjuration skill, say .1, 
they could learn conjuration spells at 1/10th the ability (and level) of a
caster with a conjuration skill of 1.0.

Skills are \emph{capped} per character level.  The formula for this is:

\[ Level \times 10 \times Class Skill Rating \]
      
For your own sanity (and balance reasons), I recommend that you use
a rating of 1.0 for class-defining skills.  Give the best swordfighting
class a `sword skill' of 1.0 and adjust other swordsmen relative to that.
You might be tempted to create something like a cavalier class with
exceptional swordfighting and grant them a 1.4 skill rating.  To illustrate
the problem this creates, let's imagine a level 10 cavalier player fighting
an equal level troll.  The troll has a moderate 0.8 clubbing skill.  Now we
have the player doing the dps of a level 14 fighter against a mob doing the
dps of a level 8.  Sure, you \emph{might} be able to balance this by making the
cavalier a paper tiger but, again, I suggest starting out conservatively.\\
    
\subsection{Advancing Skills}

Players advance skills by \emph{using} them.

NPCs share the same class skill modifiers are automatically granted max skills per their level and share the same
class modifiers as players.  
 

    

 

    
